An Automatic Verification Framework for Software Configurations

### 2016-05-10 OSDI/motiv.tex
ruzica

In this section we illustrate functionality of \app on several non-trivial configuration errors
extracted from real-world examples. Although the errors are relatively simple, we call them 
non-trivial, because the majority of existing tools, \eg, learning-based checking
tools~\cite{zhang14encore, wang04automatic}, cannot detect
these configuration errors. We investigated 
many misconfiguration problems in practice which were caused by 
these tricky configuration errors, such as those reported on StackOverflow.

\para{Example~1: Ordering error.} 
The first class of configuration errors we handle was first reported by 
Yin {\em et al.}~\cite{yin11anempirical}.
For instance, when a user configures PHP to run with the
Apache HTTP Server, this user might write, amongst others, 
the following lines in the configuration file.

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
extension = mysql.so
...
extension = recode.so
\end{lstlisting}

Such a configuration file will cause the Apache server to 
fail to start due to a segmentation fault error~\cite{yin11anempirical}. 
This is because when using PHP in Apache, the extension {\tt mysql.so} 
depends on {\tt recode.so}, and the relative ordering of two of them 
is crucial. 
We call the above example misconfiguration 
an {\em ordering error}.
Yin {\em et al.} find that ordering errors widely exist in
many system configurations, \eg, PHP and MySQL,
and typically lead to multiple system crash events;
however, there is no existing effort that can effectively solve 
or detect this problem~\cite{zhang14encore, xu15systems, xu13do}.

By using \app, the user can avoid such a configuration error.
In particular, \app is able to inform the user that {\tt recode.so} 
should appear before {\tt mysql.so}, as shown
below.

@@

\para{Example~2: Fine-grained value correlation error.} 
In a real-world misconfiguration example~\cite{correlation}, 
a user configures his MySQL as the following.

@@

This user complained that her MySQL load was very high, causing the website's
respond speed to be very slow.
In this case, {\tt key\_buffer\_size} is used by all the threads
cooperatively, while {\tt join\_buffer} and {\tt sort\_buffer} are 
created by each thread for private use; thus, the maximum amount
of used key buffer, \ie, {\tt key\_buffer\_size} should be larger than 
{\tt join|sort\_buffer\_size} * {\tt max\_connections}. 
Clearly, in the above example, it does not hold, 
so this misconfiguration causes MySQL to load very slowly.

We call the above situation a {\em fine-grained value correlation
error}, which is a more challenging case than 
{\em value correlation error} reported
by some existing investigations~\cite{yin11anempirical, zhang14encore}.
A typical value correlation error
means one entry's value should have a certain correlation with
another entry's value. For example, in MySQL,
the value of {\tt max\_connections} should be higher than
{\tt mysql.max\_persistent}. Although some work, \eg,
EnCore~\cite{zhang14encore}, can detect the normal value
correlation case, 
no existing effort can detect fine-grained value correlation
errors, which ``hide'' equation-aware correlations, as shown in
the above example.
As mentioned by Xu {\em et al.}~\cite{xu15hey},
detecting fine-grained value correlation errors 
is a much more challenging
task than the normal value correlation problem.

To the best of our knowledge, \app is the first effort that
is able to check such fine-grained value correlation problems.
If we run \app on this configuration file, \app  would return:

@@

Compared with \app, existing efforts, \eg, EnCore, cannot 
detect fine-grained value correlation errors.

\para{Example~3: Entry missing error.} 
Many critical system outages result from the fact that an important
entry was missing in the configuration file. 
We call such a problem an {\em entry missing error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world entry missing error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

When using the password policy overlay in OpenLDAP, 
users have to first include the related schema.
Leaving out the {\tt include} entry, 
as done by many users~\cite{yin11anempirical}, 
will cause failure of this LDAP. 
The user can run \app on such a misconfiguration file,
and \app would return:

>>>>

In this section we illustrate functionality of \app on several non-trivial configuration errors
extracted from real-world examples. Although the errors are relatively simple, we call them 
non-trivial, because the majority of existing tools, \eg, learning-based checking
tools~\cite{zhang14encore, wang04automatic}, cannot detect
these configuration errors. Most of the presented examples were found on StackOverflow,
a popular question and answer website for programmers. To better understand problems that users have with 
configuration files, we explored and analyzed a large number of user forums and on-line discussion sites, focused on misconfigurations.

\para{Example~1: Ordering error.} 
Ordering errors were reported by Yin {\em et al.}~\cite{yin11anempirical} and our first
example illustrates how ordering errors can cause a system to crash. When a user configures PHP 
to run with the
Apache HTTP Server, most likely the user will take some already existing configuration files and adapt them
to suite her needs. The configuration file might contain, among others, 
the following lines:
\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
extension = mysql.so
...
extension = recode.so
\end{lstlisting}

This configuration file will cause that the Apache server will
fail to start due to a segmentation fault error. 
This is because, when using PHP in Apache, the extension {\tt mysql.so} 
depends on {\tt recode.so}, and their relative ordering
is crucial. 
We call the above example of a misconfiguration file
an {\em ordering error}.
Yin {\em et al.} report that ordering errors widely exist in
many system configurations, \eg, PHP and MySQL,
and typically lead to multiple system crash events.
However, none of existing tools can effectively solve 
or detect this problem~\cite{zhang14encore, xu15systems, xu13do}.

By invoking \app, the user can detect such a configuration error.
In particular, \app reports that {\tt recode.so} 
should appear before {\tt mysql.so}, as shown
below:

@@

\para{Example~2: Fine-grained value correlation error.} 
Our next real-world misconfiguration example~\cite{correlation} comes from a discussion on StackOverflow.
The user has configured his MySQL as the following:

@@

The user the complains that her MySQL load was very high, causing the website's
respond speed to be very slow.
In this case, {\tt key\_buffer\_size} is used by all the threads
cooperatively, while {\tt join\_buffer} and {\tt sort\_buffer} are 
created by each thread for private use; thus, the maximum amount
of used key buffer, \ie, {\tt key\_buffer\_size} should be larger than 
{\tt join|sort\_buffer\_size} * {\tt max\_connections}. 
Clearly, in the above example, it does not hold, 
so this misconfiguration causes MySQL to load very slowly.

If we run \app on this configuration file, \app  would return:

@@

We can see that during the learning process we do not learn only simple statements, that compare two values, 
but we also learn more complex correlations. We call them {\em fine-grained value correlations}, and the error
a fine-grained value correlation
error. This types of errors are more sophisticated than simple value correlations that some tools can detect~\cite{yin11anempirical, zhang14encore}.
A typical value correlation error
states that one entry's value should have a certain correlation with
another entry's value. For example, in MySQL,
the value of {\tt max\_connections} should be higher than
{\tt mysql.max\_persistent}.

Our tool can learn this simple correlations as well, but learning more complex properties requires a different 
approach to the learning 
process. We need to track several variables. It is not enough just to compare them, but we need to learn, as in this particular case, which
invariant do they have to preserve. It was pointed out by Xu {\em et al.}~\cite{xu15hey} that
detecting fine-grained value correlation errors 
is a much more challenging
task than the normal value correlation problem.

To the best of our knowledge, \app is the first tool that
is able to check such fine-grained value correlation problems.

\para{Example~3: Missing entry error.} 
Many critical system outages result from the fact that an important
entry was missing in the configuration file. 
We call such a problem an {\em missing entry error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world entry missing error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

When using the password policy overlay in OpenLDAP, 
users must first include the related schema.
Leaving out the {\tt include schema/ppolicy.schema} entry, 
as done by many users~\cite{yin11anempirical}, 
causes the failure of LDAP. 
If the user runs \app on such a misconfiguration file,
\app would return:
